### 2012-04-16 trunk/tex/sigmasfr/ms.tex
checking in latex manuscript

\documentclass[apj]{emulateapj}
%\documentclass[12pt,preprint]{aastex}

\newcommand{\kms}{km~s$^{-1}$}
\newcommand{\mgii}{\textrm{Mg}~\textsc{ii}}
\newcommand{\nev}{[\textrm{Ne}~\textsc{v}]}
\newcommand{\oii}{[\textrm{O}~\textsc{ii}]}
\newcommand{\oiii}{[\textrm{O}~\textsc{iii}]}
\newcommand{\msun}{M$_{\odot}$}
\newcommand{\mstar}{M$_{*}$}
\newcommand{\units}{M$_{\odot}$~yr$^{-1}$~kpc$^{-2}$}
\newcommand{\lrest}{\lambda_{\textnormal{\scriptsize{rest}}}}
\newcommand{\lobs}{\lambda_{\textnormal{\scriptsize{obs}}}}
\newcommand{\sigmasfr}{\Sigma_{\textnormal{\scriptsize{SFR}}}}
\newcommand {\apgt} {\ {\raise-.5ex\hbox{$\buildrel>\over\sim$}}\ }
\newcommand {\aplt} {\ {\raise-.5ex\hbox{$\buildrel<\over\sim$}}\ }

\shorttitle{Compact Starbursts Revealed by HST and WISE}
\shortauthors{Diamond-Stanic et al.}

\begin{document}

%\title{HST and WISE Reveal Compact Starbursts Driving 1000
%  \kms\ Outflows}

\title{In the Crucible: Massive Compact Galaxies with High-Velocity
  Outflows and Eddington-Limited Star Formation}

\author{Aleksandar M. Diamond-Stanic\altaffilmark{1,2}, John
  Moustakas\altaffilmark{1}, Christy A. Tremonti\altaffilmark{3},
  Alison L. Coil\altaffilmark{1}, Ryan C. Hickox\altaffilmark{4}, Aday
  Robaina\altaffilmark{5}, Gregory H. Rudnick\altaffilmark{6}, \& Paul
  Sell\altaffilmark{3} }

\altaffiltext{1}{Center for Astrophysics and Space Sciences,
  University of California, San Diego, La Jolla, CA 92093, USA}
\altaffiltext{2}{Center for Galaxy Evolution Fellow}
\altaffiltext{3}{Department of Astronomy, University of
  Wisconsin-Madison, Madison, WI 53706, USA}
\altaffiltext{4}{Department of Physics and Astronomy, Dartmouth
  College, Hanover, NH 03755, USA}

\altaffiltext{5}{Institut de Ci{\'e}ncies del Cosmos, University of
  Barcelona, Barcelona, Spain}

\altaffiltext{6}{Department of Physics and Astronomy, University of
  Kansas, Lawrence, KS 66045, USA}

\begin{abstract}

Based on infrared photometry from the Wide-field Infrared Survey
Explorer (WISE) and optical morphologies from the Hubble Space
Telescope (HST), we present the discovery of compact, obscured star
formation in galaxies that exhibit $1000$~\kms\ outflows.  The implied
star formation rate (SFR) surface densities approach
$\sigmasfr=5000$~\msun~yr$^{-1}$~kpc$^{-2}$, and we argue that
feedback associated with this starburst event is a likely mechanism
for launching the high-velocity outflows.

\end{abstract}

\keywords{keywords}

\section{Introduction}

The central regions of elliptical galaxies are thought to form in
compact starbursts \citep[e.g.,][]{kor09,hop09}.  Feedback associated
with such starbursts can produce outflows driven by thermal energy
from supernova explosions \citep[e.g.,][]{che85}, stellar winds
\citep[e.g.,][]{lei92}, and momentum input from both supernova ram
pressure and radiation pressure on dust grains \citep[e.g.,][]{mur05}.
It has been argued that such feedback imposes a limit on the maximum
star-formation rate surface density ($\sigmasfr$) for starbursts
\citep[e.g.,][]{leh96,meu97,mur05,tho05} and the maximum stellar
surface density for elliptical galaxies and star clusters
\citep[e.g.,][]{hop10}.

% The relative importance of momentum injection versus thermal heating
% may vary as a function of galaxy mass \citep[e.g.,][]{hop12}, with
% thermal heating being less efficient in galaxies with larger gas
% densities and shorter cooling times.  Massive star clusters with large
% gas surface densities are a potential launching point for
% galactic-scale outflows driven by radiation pressure
% \citep[e.g.,][]{mur11}.

Galactic winds are ubiquitous in star-forming galaxies at all
redshifts and generally exhibit outflow velocities in the
100--500~\kms\ range
\citep[e.g.,][]{hec00,sha03,mar05,rup05,wei09,rub10,ste10}.  Thus, the
discovery of $|v|>1000$~\kms\ outflows in a sample of massive
($\textnormal{\mstar}\approx10^{11}$~\msun) post-starburst galaxies at
$z\sim0.6$ by \citet{tre07} suggested that a more energetic source,
such as feedback from an accreting supermassive black hole
\citep[e.g.,][]{sil98,dim05,deb12}, could have been responsible for
launching the winds.

However, it also plausible that feedback from a compact starburst
could expel gas with such large velocities.  Indeed, there is evidence
for a positive correlation between outflow velocity and starburst
luminosity \citep[e.g.,][]{mar05,rup05,tre07}, albeit with significant
scatter.  Furthermore, \citet{hec11} recently found outflows with
maximum velocities reaching $-1500$~\kms\ in a sample of local
starbursts with compact nuclei, and argued that such velocities could
be explained by feedback from massive stars.

% $|v|\gtrsim1000$~\kms\

In this Letter, we present results for a sample of massive galaxies at
$z\sim0.6$ that exhibit $|v|\apgt1000$~\kms\ outflows, expanding on
the sample from \citet{tre07}.  We seek to test whether the energetic
outflows in these galaxies could have been driven by feedback from
starbursts with very large star-formation rate (SFR) surface densities
($\sigmasfr$).  Our analysis combines galaxy sizes obtained with the
Hubble Space Telescope (HST) with star-formation rates and stellar
masses estimated from WISE, Spitzer, SDSS, and GALEX photometry.

% For example, there is evidence for a positive correlation between
% outflow velocity and starburst luminosity \citep[e.g.,][]{mar05},
% albeit with significant scatter.  Furthermore, \citet{hec11} recently
% found outflows reaching velocities of $-1500$~\kms\ in a sample of
% local starbursts with compact nuclei, and argued their higher
% velocities could be explained by feedback exerted by a large number of
% massive stars that are confined to a small radius.

% [which has been invoked theoretically as
%  a mechanism to quench star formation in galaxies,

% Winds are also common in active galactic nuclei (AGNs) with velocities
% in the 1000--20,000~\kms range \citep[e.g., ][]{cre03}, although it is
% unclear whether AGN feedback plays an important role in regulating
% star formation in galaxies.  There is clear observational evidence
% that active galactic nuclei (AGNs) produce outflows
% \citep[e.g.,][]{cre03} and that the outflowing gas can extend to
% $>$kpc scales \citep[e.g.,][]{rup11}, it is unclear whether such
% feedback plays an important role globally in quenching star formation.

% Feedback from massive stars and active black holes has important
% implications for galaxy evolution (e.g., luminosity function, star
% formation, galaxy color bi-modality, mass-metallicity relation,
% enrichment of IGM).

% Feedback from accreting supermassive black holes has also been invoked
% theoretically as a mechanism for quenching star formation in galaxies
% \citep[e.g.,][]{sil98,gra04,dim05}.  There is clear observational
% evidence that active galactic nuclei (AGNs) produce outflows
% \citep[e.g.,][]{cre03} and that the outflowing gas can extend to
% $>$kpc scales \citep[e.g.,][]{rup11}, but it is unclear whether such
% feedback plays an important role globally in quenching star formation
% and shaping the galaxy mass function.

% and optical spectra ($\lrest=2500$-5200~\AA)

%  The spectral region surrounding the
%  \mgii~$\lambda\lambda2796,2803$ absorption alines, which are used to
%  measure outflow velocities, is highlighted in both wavelength and
%  velocity space.

%  and are scaled to match the WISE channel 2 photometry ($\lobs=4.6$~$\mu$m)
% (\mgii~$\lambda2796$ is plotted in blue and \mgii~$\lambda2803$ is plotted in red)

\section{Analysis}

\subsection{HST/WFC3 Imaging}

Our sample derives from 29 galaxies targeted for HST/WFC3 imaging in
programs 12019 and 12272 (PI: C. Tremonti).  Using the F814W filter on
the UVIS channel, which has $0.04\arcsec$ pixels and
$\textnormal{FWHM}=0.074\arcsec$, we obtained $4\times10$~min
exposures in a single orbit for each galaxy.  The dithered images were
processed with
MultiDrizzle\footnote{http://stsdas.stsci.edu/multidrizzle/} to
produce science mosaics with $0.02\arcsec$ pixels.  For each galaxy,
we use GALFIT \citep{pen02,pen10} to model the two-dimensional surface
brightness profile with a single Sersic component, using stars in the
images to construct the model point-spread function (PSF).  The Sersic
index (n) and effective radius ($r_e$) are free parameters in the
model.  In cases where the best-fit model returns $n>4$, we also fit
an $n=4$ de Vaucouleurs model \citep{dev48}, yielding a larger $r_e$
value (due to the covariance between $n$ and $r_e$), and we use these
these larger effective radii in our analysis.

For this paper, we are most interested in the galaxies with the
largest $\sigmasfr$ values, which also have the smallest effective
radii.  We show the HST images for three high-$\sigmasfr$ galaxies in
Figure~\ref{fig:seds}.  In all three cases, the single-component model
GALFIT accounts for $>85\%$ of the total flux.  The residuals show
diffuse emission that is consistent with these systems being
late-stage galaxy mergers after final coalescence, although we defer a
detailed comparison of the observed morphologies with expectations
from merger simulations to future work.

For the most compact galaxy (J0905+5759, $r_e=0.013\arcsec$ or
100~pc), we also show the observed one-dimensional surface brightness
profile in the bottom-right panel of Figure~\ref{fig:profile}.  We
compare to the profiles of six stars in the same image, the best-fit
de Vaucouleurs model, and a de Vaucoulerus model with a larger
effective radius ($r_e=0.04\arcsec$, the native WFC3/UVIS pixel size,
which corresponds to a physical scale of 290~pc).  This comparison
illustrates that this galaxy, while only marginally resolved with an
effective radius that is $\sim20\%$ of the image FWHM, is clearly more
extended than a point source.

For such a compact source, there is significant uncertainty in our
$r_e$ measurement given uncertainties in the model PSF.  To quantify
this uncertainty, we used TinyTim to generate a model PSF that is
artificially narrower than the stars in the image (convolving the
output of tiny2 with a $\textnormal{FWHM}=0.04\arcsec$ Gaussian,
whereas the image FWHM is $0.074\arcsec$) and found that this
increased the effective radius in the GALFIT model by a factor of two.
We also fit a two-component PSF+Sersic model, but found that the
Sersic component dominates the fit, yielding a similar effective
radius.  Furthermore, the spectrum of the galaxy shows no evidence for
an AGN contribution to the $I$-band continuum (see
Figure~\ref{fig:spectra}), so there is no clear motivation for
including an unresolved, point-source component in the model.  We
conclude that this galaxy is quite compact and that our $r_e$ estimate
of 100~pc, while uncertain, is likely accurate within a factor of two.

\begin{figure}[!t]
%\includegraphics[angle=90,scale=0.68]{sigmasfr_datafigure.ps}
\includegraphics[angle=0,scale=0.41]{3seds.ps}
\caption{Left: Observed UV--IR SEDs ($\lrest=0.1$--15~$\mu$m) for
  three galaxies with large SFR surface densities
  ($\sigmasfr>3000$~\units).  We show stellar population fits to the
  $\lrest=0.1$--3~$\mu$m emission (black solid line) and three
  templates for dust emission \citep[M82 starburst, Arp 220 starburst,
    and obscured quasar;][]{pol07}.  The starburst templates provide
  reasonable fits to the $\lobs=12$ and 22~$\mu$m WISE photometry,
  while the quasar template does not.  Right: HST/WFC3 F814W images
  (probing $\lrest\approx5000$~\AA) showing that these galaxies are
  dominated by a compact nucleus.}
\label{fig:seds}
\end{figure}
% scaled to the WISE $\lobs=4.6~\mu$m flux.

\begin{figure}[!t]
\includegraphics[angle=0,scale=0.41]{profile.ps}
\label{fig:profile}
\caption{One-dimensional surface brightness profile for J0905+5759,
  which has the smallest effective radius in the sample.  The observed
  profile is shown as the solid black line, and the profiles of six
  stars in the same image are shown in blue.  The best-fit de
  Vaucouleurs profile with $r_e=100$~pc is shown as a dashed red line,
  and for comparison a broader profile with $r_e=0.04\arcsec=290$~pc
  is shown as the dotted green line.  This galaxy is quite compact,
  but is more extended than a point source.}
\end{figure}

\subsection{Photometry, Stellar Masses, and Star-formation rates}

We gathered photometry from the All-Sky Release of the Wide-field
Infrared Survey Explorer \citep[WISE,][]{wri10}, the Seventh Data
Release of the Sloan Digital Sky Survey \citep[SDSS,][]{aba09}, and
General Release 6 from the Galaxy Evolution Explorer (GALEX).  We also
obtained $5\times30$~sec dithered exposures in the 3.6~$\mu$m and
4.5~$\mu$m for all sources with the Infrared Array Camera
\citep{faz04} on the Spitzer Space Telescope \citep{wer04} as part of
General Observer program 60145 (PI: C. Tremonti).  We used the
post--basic calibrated data to perform aperture photometry on all
sources, and we also used the
APEX\footnote{http://irsa.ipac.caltech.edu/data/SPITZER/docs/dataanalysistools/tools/mopex/}
point-source extraction software for photometry of sources in crowded
fields.  All the photometry was corrected for Galactic extinction
based on the \citet{sch98} dust maps.  We show spectral energy
distributions (SEDs) for the three high-$\sigmasfr$ galaxies in
Figure~\ref{fig:seds}.

We use the method of \citet{mou11} to derive stellar masses and
UV-based SFRs from fits to the 0.1--5~$\mu$m SEDs.  We also estimate
IR-based SFRs from the 12~$\mu$m and 22~$\mu$m WISE photometry.

Comment on agreement between SFR indicators.

\begin{figure}[!t]
\includegraphics[angle=0,scale=0.41]{spectra.ps}
\caption{Spectra covering $\lrest=2500$--5200~\AA\ for the three
  galaxies shown in Figure~\ref{fig:seds}.  These spectra are
  dominated by light from the young stellar population and exhibit
  \oii~$\lambda3727$, H$\beta$~$\lambda4861$, and \oiii~$\lambda5007$
  emission lines.  The panels on the right zoom on the
  \mgii~$\lambda\lambda2796,2803$ spectral region in both wavelength
  and velocity space.  The spectrum in the bottom panel has sufficient
  spectral resolution ($\textnormal{FWHM}\approx8$~\kms) to resolve
  the intrinsic shape of the absorption-line profile, revealing that
  the gas near the centroid velocity has unity covering factor.}
\label{fig:spectra}
\end{figure}

\begin{figure*}[!t]
\begin{center}
\includegraphics[angle=0,scale=.8]{sigmasfr.ps} %\\
%\includegraphics[angle=0,scale=.7]{sizemass_sigmasfr.ps}
%\caption{$\sigmasfr$--mass and size--mass relationships.}
\caption{SFR surface densities and stellar masses for the HST--WISE
  sample described in this paper (black circles, with symbol size
  proportional to outflow velocity), along with samples of gas-rich
  mergers (orange squares) and star-forming galaxies (shown with 68\%
  contours).  We mark the empirical threshold for launching winds
  \citep[dotted line, $\sigmasfr=0.1$~\units;][]{hec02}, the
  90th-percentile starburst intensity limit from \citet{meu97} (dashed
  line, $\sigmasfr=45$~\units), and the Eddington limit from radiation
  pressure on dust grains \citep[solid line,
    $\sigmasfr\approx2000$~\units;][]{mur05,tho05,hop10}.  Our
  HST--WISE sample overlaps with the region characterized by gas-rich
  mergers, and extends to very large SFR surface densities near the
  Eddington limit.  The sample of gas-rich mergers includes ULIRGs
  \citep{vei06}, Lyman break analogs with dominant central objects
  \citep{ove09}, and the local compact starburst Arp 220 \citep{sco97,
    ken98, rod08}.  The samples of star-forming galaxies are drawn
  from \citet{wuy11} at $0.02<z<0.2$ (solid black contour),
  $0.5<z<1.5$ (dashed red countour), and $1.5<z<2.5$ (dotted purple
  contour), and from \citet{law12} at $2.5<z<3.6$ (dash-dotted blue
  contour). }
\label{fig:sigmasfr}
\end{center}
\end{figure*}

\subsection{Optical Spectroscopy}

We present low-resolution $R=1000-3000$ spectroscopy from MMT/Blue
Channel and SDSS (J1506+5402), Magellan/MagE (J1341-0321), and
Keck/LRIS (J0905+5759) in Figure~\ref{fig:spectra.}.  We also
highlight the \mgii~$\lambda\lambda2796,2803$ absorption lines, which
are used use to measure outflow velocities.

For J0905+5759 we show the \mgii\ absorption-line profiles from
$R\approx37,000$ spectra obtained with Keck/HIRES.  The oscillator
strength of \mgii~$\lambda$2796 is twice that of \mgii~$\lambda$2803,
but the equivalent widths of the two lines and the shapes of the line
profiles are quite similar, indicating that both lines are optically
thick.  The fact that the absorption line profile approaches zero near
the velocity centroid $v=-2470$~\kms\ indicates that the outflowing
gas covers the entire continuum source at these velocities.

%\section{Analysis}
%
%\subsection{Star-formation Rates and Stellar Masses}
%
%\subsection{Morphologies and Sizes}
%
%\subsection{Outflow Kinematics and Covering Factors}

% \begin{figure*}[!t]
% \begin{center}
% \includegraphics[angle=0,scale=.8]{sizemass_sigmasfr.ps}
% \caption{size--mass relationship.}
% \label{fig:sizemass}
% \end{center}
% \end{figure*}

% %\begin{center}
% \begin{deluxetable}{ccccccc}
% \tabletypesize{\scriptsize}
% 
% \tablecaption{Galaxy Properties\label{tab:properties}}
% \tablewidth{0pt}
% 
% \tablehead{ \colhead{Galaxy Name} & \colhead{z} & \colhead{$M_{*}$} & \colhead{$v$} & \colhead{SFR} & \colhead{$r_e$} & \colhead{$\sigmasfr$} }
% 
% \startdata
% J150636.29+540220.7   & 0.608 & 11.01 & -1211 &  621 & 0.13 & 5851 \\
% J090523.60+575912.4   & 0.712 & 10.62 & -2470 & 344 & 0.10 & 5478 \\ 
% %J094417.84+093019.3   & 0.514 & 10.73 & -1778 & $<43.22$ & 161 & 0.15 & 1139 \\
% J134136.79$-$032125.2   & 0.658 & 10.61 & -875  & 490 & 0.16 & 3058 \\ 
% % J1506 has Lx=42.66
% \enddata
% 
% %\tablecomments{}
% %\tablenotetext{a}{Star-formation rate based on \citet{ken98}.  No extinction correction has been applied.}
% 
% \end{deluxetable}
% %\end{center}

\section{Discussion}

The compact sizes ($r_e\approx100$~pc) and large star formation rates
($\textnormal{SFR}\approx500$~\msun) for the galaxies shown in
Figure~\ref{fig:seds} imply extremely large star-formation rate
surface densities ($\sigmasfr\approx5000$~\units).  To place these
galaxies in context, we plot their $\sigmasfr$ values as a function of
stellar mass in Figure~\ref{fig:sigmasfr}.  We also include
measurements of late-stage, gas-rich mergers at $z<0.3$ from
\citet{vei06}, star-forming galaxies at $z<0.3$ from \citet{ove09},
and star-forming galaxies at $z=1.5$--3.6 from \citet{law12}.

Points to make:
\begin{itemize}
\item \citet{ken98}, the standard reference for the relationship
  between SFR and gas surface density, includes Arp 200 with
  $\sigmasfr\approx10^3$~\units\ and
  $\Sigma_{H_2}\approx5\times10^{10}$~\msun~kpc$^{-2}$.  
\item The star clusters considered by \citet{meu97} extend into this
  $\sigmasfr>10^{3}$~\units\ range.

\end{itemize}

% We show relationship between effective radius and stellar mass for our
% WISE-HST sample in bottom panel of Figure~\ref{fig:sizemass}.  Here,
% we include the gas-rich mergers from \citet{vei06} as well as
% $z\sim2.3$ compact massive galaxies from \citet{van08}.  We also show
% the size--mass relationship for elliptical galaxies at $z\sim0.1$ from
% \citet{she03}, and the average size evolution of spheroidal ($n>2.5$)
% galaxies from $z=2$ to $z=0.1$ from \citet{tru07}.

%\subsection{A Maximum Star-Formation Rate Surface Density?}

\subsection{Momentum Flux from Supernovae and Stellar Winds}

\citet{meu97} argued for a limit near 45~\units (somewhat smaller for
a Chabrier IMF) based on the surface brightness of starbursts measured
using rest-frame UV, H$\alpha$, far-IR, and radio continuum emission,
such that only 10\% of starbursts exceeded this limit.  They argued
that the momentum flux from supernovae and stellar winds expected at
this threshold matches the mean central pressure measured by
\citet{hec90} for starburst-driven winds.

There are also relevant arguments about ram pressure driving of
high-velocity winds from \citet{hec11}.

\subsection{The Eddington Limit from Radiation Pressure on Dust Grains}

In models where gaseous outflows are driven by radiation pressure from
massive stars, estimates of the maximum $\sigmasfr$ where the outward
force from radiation exceeds the inward force of gravity are in the
$\sigmasfr\sim10^3$~\units\ range \citep[e.g.,][]{tho05,hop10}.

% Our galaxies exceed the \citet{meu97} limit but are roughly in line
% with the limit from radiation pressure.

The relative importance of momentum injection versus thermal heating
may vary as a function of galaxy mass \citep[e.g.,][]{hop12}, with
thermal heating being less efficient in galaxies with larger gas
densities and shorter cooling times.  Regarding momentum injection,
\citet{sha11} have argued that radiation pressure becomes more
important than ram pressure at large SFRs.  Massive star clusters with
large gas surface densities are the ideal launching point for
galactic-scale outflows driven by radiation pressure
\citep[e.g.,][]{mur11}, and the outflow velocities are expected to
scale with escape velocity of the most massive star clusters in a
galaxy.

\subsection{Outflows at the Escape Velocity}

The observed $\gtrsim1000$~\kms\ outflow velocities we observe correspond
approximately to the escape velocities for compact regions (i.e.,
massive star clusters) with the sizes and masses that we've measured.

\begin{equation}
v_{esc} = \sqrt{2GM_{cl}/r_{cl}} = 2074~\textnormal{km~s}^{-1} \\%\left({v_c\over200}\right)^{5/2} 
\left({M_{cl}\over10^{11} M_{\odot}}\right)^{1/2} 
\left({r_{cl}\over200~\textnormal{pc}}\right)^{-1/2}
\label{eq:vesc}
\end{equation}

% This is in agreement with the predicitons from radiation pressure
% launching winds from the most massive star clusters in a galaxy.

Question of whether other observables are plausible.  
\begin{itemize}
\item Expected velocity dispersion.  Some results have suggested that
  turbulence becomes large at high surface densities, and therefore
  the velocity dispersion should be large.  
\item The dynamical time for the starburst.  If gas is consumed with
  an efficiency of a few \% per dynamical time, how long do we expect
  the starburst to last?
\end{itemize}

Other points: 
\begin{itemize}
\item Is this mechanism viable for the whole sample?  We see large
  outflow velocities in rest of the sample, where the galaxies are
  more ``post'' starburst.  This fits in a scenario where the genuine
  post-starbursts are further along in the evolution after the outflow
  quenches star formation.
\item Will these evolve into ellipticals, or do they have enough gas
  to reform a disk?  If they do quench, how difficult would it be to
  grow in size to match the local size--mass relationship?  Could
  their descendants be the handful of local galaxies that don't lie on
  the size--mass relationship?
\item Is there disagreement between $L_{IR}+L_{UV}$ and H$\alpha$
  SFRs?  \citet{hec11} pointed out that this could be a sign of
  escaping ionizing photons.
\end{itemize}

\subsection{Limits on Ongoing AGN Activity}

No evidence in spectra for unobscured AGN component.  J1506+5402 has
X-ray and \nev.

\subsection{Placing These Galaxies in Context}

They are rare and thus represent an unusual or short-lived phase.

Connection to models.  How such large SFRs given the powerful wind?

Connection to ellipticals.  Conceivable that they could grown in size
sufficiently by $z=0$?

\acknowledgments

We acknowledge useful discussions with and assistance from several
people.  AMD acknowledges support from the Southern California Center
for Galaxy Evolution, a multi-campus research program funded by the
University of California Office of Research.

% James Bullock, Xavier Prochaska, Kate Rubin, Art Wolfe, Alex Mendez

\begin{thebibliography}{}
\bibitem[Abazajian et al.(2009)]{aba09} Abazajian, K.~N.,
  Adelman-McCarthy, J.~K., Ag{\"u}eros, M.~A., et al.\ 2009, \apjs,
  182, 543
% \bibitem[Baldwin et al.(1981)]{bal81} Baldwin, J.~A., Phillips, M.~M.,
%   \& Terlevich, R.\ 1981, \pasp, 93, 5
% \bibitem[Barton \& Cooke(2009)]{bar09} Barton, E.~J., \& Cooke,
%   J.\ 2009, \aj, 138, 1817
% \bibitem[Bordoloi et al.(2011)]{bor11} Bordoloi, R., et al.\ 2011,
%   arXiv:1106.0616
% \bibitem[Buitrago et al.(2008)]{bui08} Buitrago, F., Trujillo, I.,
%   Conselice, C.~J., Bouwens, R.~J., Dickinson, M., \& Yan, H.\ 2008,
%   \apjl, 687, L61
% \bibitem[Coil et al.(2011)]{coi11} Coil, A.~L., Weiner, B.~J., Holz,
%   D.~E., et al.\ 2011, \apj, 743, 46
% \bibitem[Chen et al.(2010)]{hchen10} Chen, H.-W., Wild, V., Tinker,
%   J.~L., Gauthier, J.-R., Helsby, J.~E., Shectman, S.~A., \& Thompson,
%   I.~B.\ 2010, \apjl, 724, L176
% \bibitem[Chen et al.(2010b)]{ychen10} Chen, Y.-M., Tremonti, C.~A.,
%   Heckman, T.~M., Kauffmann, G., Weiner, B.~J., Brinchmann, J., \&
%   Wang, J.\ 2010, \aj, 140, 445
\bibitem[Chevalier \& Clegg(1985)]{che85} Chevalier, R.~A., \& Clegg,
  A.~W.\ 1985, \nat, 317, 44
% \bibitem[Coil et al.(2011)]{coi11} Coil, A.~L., Weiner, B.~J., Holz,
%   D.~E., Cooper, M.~C., Yan, R., \& Aird, J.\ 2011, arXiv:1104.0681
\bibitem[Crenshaw et al.(2003)]{cre03} Crenshaw, D.~M., Kraemer,
  S.~B., \& George, I.~M.\ 2003, \araa, 41, 117
% \bibitem[Croton et al.(2006)]{cro06} Croton, D.~J., et al.\ 2006,
%   \mnras, 365, 11
% \bibitem[Dekel \& Birnboim(2006)]{dek06} Dekel, A., \& Birnboim,
%   Y.\ 2006, \mnras, 368, 2
\bibitem[de Vaucouleurs(1948)]{dev48} de Vaucouleurs, G.\ 1948,
  Annales d'Astrophysique, 11, 247
\bibitem[Debuhr et al.(2012)]{deb12} Debuhr, J., Quataert, E., \& Ma,
  C.-P.\ 2012, \mnras, 420, 2221
\bibitem[Di Matteo et al.(2005)]{dim05} Di Matteo, T., Springel, V.,
  \& Hernquist, L.\ 2005, \nat, 433, 604
\bibitem[Fazio et al.(2004)]{faz04} Fazio, G.~G., Hora, J.~L., Allen,
  L.~E., et al.\ 2004, \apjs, 154, 10
\bibitem[Granato et al.(2004)]{gra04} Granato, G.~L., De Zotti, G.,
  Silva, L., Bressan, A., \& Danese, L.\ 2004, \apj, 600, 580
\bibitem[Heckman et al.(1990)]{hec90} Heckman, T.~M., Armus, L., \&
  Miley, G.~K.\ 1990, \apjs, 74, 833
\bibitem[Heckman et al.(2000)]{hec00} Heckman, T.~M., Lehnert, M.~D.,
  Strickland, D.~K., \& Armus, L.\ 2000, \apjs, 129, 493
\bibitem[Heckman(2002)]{hec02} Heckman, T.~M.\ 2002, Extragalactic Gas
  at Low Redshift, 254, 292
\bibitem[Heckman et al.(2011)]{hec11} Heckman, T.~M., et al.\ 2011,
  \apj, 730, 5
% \bibitem[Hopkins et al.(2009)]{hop09} Hopkins, P.~F., Bundy, K.,
%   Murray, N., Quataert, E., Lauer, T.~R., \& Ma, C.-P.\ 2009, \mnras,
%   398, 898
%\bibitem[Hopkins et al.(2009)]{hop09} Hopkins, P.~F., Hernquist, L.,
%  Cox, T.~J., Keres, D., \& Wuyts, S.\ 2009, \apj, 691, 1424
\bibitem[Hopkins et al.(2009)]{hop09} Hopkins, P.~F., Cox, T.~J.,
  Dutta, S.~N., et al.\ 2009, \apjs, 181, 135
% \bibitem[Hopkins et al.(2010a)]{hop10a} Hopkins, P.~F.,
%   Bundy, K., Hernquist, L., Wuyts, S., \& Cox, T.~J.\ 2010, \mnras,
%   401, 1099
\bibitem[Hopkins et al.(2010)]{hop10} Hopkins, P.~F., Murray, N.,
  Quataert, E., \& Thompson, T.~A.\ 2010, \mnras, 401, L19
% %\bibitem[Hopkins et al.(2011)]{hop11} Hopkins, P.~F., Quataert, E., \&
% %  Murray, N.\ 2011, arXiv:1101.4940
% \bibitem[Hopkins et al.(2011)]{hop11} Hopkins, P.~F., Quataert, E., \&
%   Murray, N.\ 2011, arXiv:1110.4638
\bibitem[Hopkins et al.(2012)]{hop12} Hopkins, P.~F., Quataert, E., \&
  Murray, N.\ 2012, \mnras, arXiv:1110.4638
% \bibitem[Kacprzak et al.(2010)]{kac10} Kacprzak, G.~G., Churchill,
%   C.~W., Ceverino, D., Steidel, C.~C., Klypin, A., \& Murphy,
%   M.~T.\ 2010, \apj, 711, 533
% \bibitem[Kacprzak et al.(2011a)]{kac11a} Kacprzak, G.~G., Churchill,
%   C.~W., Barton, E.~J., \& Cooke, J.\ 2011, \apj, 733, 105
% \bibitem[Kacprzak et al.(2011b)]{kac11b} Kacprzak, G.~G., Churchill,
%   C.~W., Evans, J.~L., Murphy, M.~T., \& Steidel, C.~C.\ 2011,
%   arXiv:1106.3068
% \bibitem[Kauffmann et al.(2003)]{kau03} Kauffmann, G., Heckman, T.~M.,
%   Tremonti, C., et al.\ 2003, \mnras, 346, 1055
% \bibitem[Kennicutt(1998)]{ken98} Kennicutt, R.~C., Jr.\ 1998, \araa,
%   36, 189
\bibitem[Kennicutt(1998)]{ken98} Kennicutt, R.~C., Jr.\ 1998, \apj,
  498, 541
% \bibitem[Kere{\v s} et al.(2005)]{ker05} Kere{\v s}, D., Katz, N.,
%   Weinberg, D.~H., \& Dav{\'e}, R.\ 2005, \mnras, 363, 2
% \bibitem[Kere{\v s} et al.(2009)]{ker09} Kere{\v s}, D., Katz, N.,
%   Dav{\'e}, R., Fardal, M., \& Weinberg, D.~H.\ 2009, \mnras, 396,
%   2332
% \bibitem[Kobulnicky \& Kewley(2004)]{kob04} Kobulnicky, H.~A., \&
%   Kewley, L.~J.\ 2004, \apj, 617, 240
\bibitem[Kormendy et al.(2009)]{kor09} Kormendy, J., Fisher, D.~B.,
  Cornell, M.~E., \& Bender, R.\ 2009, \apjs, 182, 216
% \bibitem[Kriek et al.(2010)]{kri10} Kriek, M., et al.\ 2010, \apjl,
%   722, L64
\bibitem[Law et al.(2012)]{law12} Law, D.~R., Steidel, C.~C., Shapley,
  A.~E., et al.\ 2012, \apj, 745, 85
\bibitem[Lehnert \& Heckman(1996)]{leh96} Lehnert, M.~D., \& Heckman,
  T.~M.\ 1996, \apj, 472, 546
% \bibitem[Lehnert \& Heckman(1996)]{leh96} Lehnert, M.~D., \& Heckman,
%   T.~M.\ 1996, \apj, 462, 651
\bibitem[Leitherer et al.(1992)]{lei92} Leitherer, C., Robert, C., \&
  Drissen, L.\ 1992, \apj, 401, 596
\bibitem[Martin(2005)]{mar05} Martin, C.~L.\ 2005, \apj, 621, 227
% \bibitem[Martin(2006)]{mar06} Martin, C.~L.\ 2006, \apj, 647, 222
% \bibitem[Martin \& Bouch{\'e}(2009)]{mar09} Martin, C.~L., \&
%   Bouch{\'e}, N.\ 2009, \apj, 703, 1394
% \bibitem[McIntosh et al.(2005)]{mci05} McIntosh, D.~H., et al.\ 2005,
%   \apj, 632, 191
\bibitem[Meurer et al.(1997)]{meu97} Meurer, G.~R., Heckman, T.~M.,
  Lehnert, M.~D., Leitherer, C., \& Lowenthal, J.\ 1997, \aj, 114, 54
% \bibitem[Morrissey et al.(2007)]{mor07} Morrissey, P., Conrow, T.,
%   Barlow, T.~A., et al.\ 2007, \apjs, 173, 682
\bibitem[Moustakas et al.(2011)]{mou11} Moustakas, J., Zaritsky, D.,
  Brown, M., et al.\ 2011, arXiv:1112.3300
\bibitem[Murray et al.(2005)]{mur05} Murray, N., Quataert, E., \&
  Thompson, T.~A.\ 2005, \apj, 618, 569
\bibitem[Murray et al.(2011)]{mur11} Murray, N., M{\'e}nard, B., \&
  Thompson, T.~A.\ 2011, \apj, 735, 66
% \bibitem[Nath \& Silk(2009)]{nat09} Nath, B.~B., \& Silk, J.\ 2009,
%   \mnras, 396, L90
% \bibitem[Nestor et al.(2011)]{nes11} Nestor, D.~B., Johnson, B.~D.,
%   Wild, V., M{\'e}nard, B., Turnshek, D.~A., Rao, S., \& Pettini,
%   M.\ 2011, \mnras, 412, 1559
% \bibitem[Oke(1990)]{oke90} Oke, J.~B.\ 1990, \aj, 99, 1621
% \bibitem[Oke et al.(1995)]{oke95} Oke, J.~B., et al.\ 1995, \pasp,
%   107, 375
% \bibitem[Oppenheimer \& Dav{\'e}(2006)]{opp06} Oppenheimer, B.~D., \&
%   Dav{\'e}, R.\ 2006, \mnras, 373, 1265
% \bibitem[Osterbrock(1989)]{ost89} Osterbrock, D.~E.\ 1989,
%   Astrophysics of Gaseous Nebulae and Active Galactic Nuclei (Mill
%   Valley, CA: University Science Books) 
\bibitem[Overzier et al.(2009)]{ove09} Overzier, R.~A., Heckman,
  T.~M., Tremonti, C., et al.\ 2009, \apj, 706, 203
% \bibitem[Pagel et al.(1979)]{pag79} Pagel, B.~E.~J., Edmunds, M.~G.,
%   Blackwell, D.~E., Chun, M.~S., \& Smith, G.\ 1979, \mnras, 189, 95
\bibitem[Peng et al.(2002)]{pen02} Peng, C.~Y., Ho, L.~C., Impey,
  C.~D., \& Rix, H.-W.\ 2002, \aj, 124, 266
\bibitem[Peng et al.(2010)]{pen10} Peng, C.~Y., Ho, L.~C., Impey,
  C.~D., \& Rix, H.-W.\ 2010, \aj, 139, 2097
% \bibitem[Pettini et al.(2001)]{pet01} Pettini, M., Shapley, A.~E.,
%   Steidel, C.~C., Cuby, J.-G., Dickinson, M., Moorwood, A.~F.~M.,
%   Adelberger, K.~L., \& Giavalisco, M.\ 2001, \apj, 554, 981
\bibitem[Polletta et al.(2007)]{pol07} Polletta, M., Tajer, M.,
  Maraschi, L., et al.\ 2007, \apj, 663, 81
% \bibitem[Prochaska \& Hennawi(2009)]{pro09} Prochaska, J.~X., \&
%   Hennawi, J.~F.\ 2009, \apj, 690, 1558
% \bibitem[Prochaska et al.(2011a)]{pro11a} Prochaska, J.~X., Kasen, D.,
%   \& Rubin, K.\ 2011, arXiv:1102.3444
% \bibitem[Prochaska et al.(2011b)]{pro11b} Prochaska,
%   J.~X., Weiner, B., Chen, H.~-., Mulchaey, J.~S., \& Cooksey,
%   K.~L.\ 2011, arXiv:1103.1891
\bibitem[Rodr{\'{\i}}guez Zaur{\'{\i}}n et al.(2008)]{rod08}
  Rodr{\'{\i}}guez Zaur{\'{\i}}n, J., Tadhunter, C.~N., \&
  Gonz{\'a}lez Delgado, R.~M.\ 2008, \mnras, 384, 875
% \bibitem[Rubin et al.(2010a)]{rub10a} Rubin, K.~H.~R., Prochaska, J.~X.,
%   Koo, D.~C., Phillips, A.~C., \& Weiner, B.~J.\ 2010, \apj, 712, 574
\bibitem[Rubin et al.(2010)]{rub10} Rubin, K.~H.~R., Weiner, B.~J.,
  Koo, D.~C., Martin, C.~L., Prochaska, J.~X., Coil, A.~L., \& Newman,
  J.~A.\ 2010, \apj, 719, 1503
% \bibitem[Rubin et al.(2011)]{rub11} Rubin, K.~H.~R., Prochaska, J.~X.,
%   M{\'e}nard, B., Murray, N., Kasen, D., Koo, D.~C., \& Phillips,
%   A.~C.\ 2011, \apj, 728, 55
\bibitem[Rupke et al.(2005)]{rup05} Rupke, D.~S., Veilleux, S., \&
  Sanders, D.~B.\ 2005, \apjs, 160, 115 % analysis and discussion
% \bibitem[Rupke \& Veilleux(2011)]{rup11} Rupke, D.~S.~N., \& Veilleux,
%   S.\ 2011, \apjl, 729, L27
% \bibitem[Sato et al.(2009)]{sat09} Sato, T., Martin, C.~L., Noeske,
%   K.~G., Koo, D.~C., \& Lotz, J.~M.\ 2009, \apj, 696, 214
\bibitem[Schlegel et al.(1998)]{sch98} Schlegel, D.~J., Finkbeiner,
  D.~P., \& Davis, M.\ 1998, \apj, 500, 525
% \bibitem[Schwartz \& Martin(2004)]{sch04} Schwartz, C.~M., \& Martin,
%   C.~L.\ 2004, \apj, 610, 201
\bibitem[Scoville et al.(1997)]{sco97} Scoville, N.~Z., Yun, M.~S., \&
  Bryant, P.~M.\ 1997, \apj, 484, 702
\bibitem[Sharma \& Nath(2011)]{sha11} Sharma, M., \& Nath,
  B.~B.\ 2011, arXiv:1112.3447
\bibitem[Shapley et al.(2003)]{sha03} Shapley, A.~E., Steidel, C.~C.,
  Pettini, M., \& Adelberger, K.~L.\ 2003, \apj, 588, 65
\bibitem[Shen et al.(2003)]{she03} Shen, S., Mo, H.~J., White,
  S.~D.~M., et al.\ 2003, \mnras, 343, 978
\bibitem[Silk \& Rees(1998)]{sil98} Silk, J., \& Rees, M.~J.\ 1998,
  \aap, 331, L1
% \bibitem[Springel \& Hernquist(2003)]{spr03} Springel, V., \&
%   Hernquist, L.\ 2003, \mnras, 339, 312
% \bibitem[Steidel et al.(2002)]{ste02} Steidel, C.~C., Kollmeier,
%   J.~A., Shapley, A.~E., Churchill, C.~W., Dickinson, M., \& Pettini,
%   M.\ 2002, \apj, 570, 526
\bibitem[Steidel et al.(2010)]{ste10} Steidel, C.~C., Erb, D.~K.,
  Shapley, A.~E., Pettini, M., Reddy, N., Bogosavljevi{\'c}, M.,
  Rudie, G.~C., \& Rakic, O.\ 2010, \apj, 717, 289
% \bibitem[Stewart et al.(2011a)]{ste11a} Stewart, K.~R., Kaufmann, T.,
%   Bullock, J.~S., Barton, E.~J., Maller, A.~H., Diemand, J., \&
%   Wadsley, J.\ 2011, \apjl, 735, L1
% \bibitem[Stewart et al.(2011b)]{ste11b} Stewart, K.~R., Kaufmann, T.,
%   Bullock, J.~S., Barton, E.~J., Maller, A.~H., Diemand, J., \&
%   Wadsley, J.\ 2011, arXiv:1103.4388
% \bibitem[Strickland et al.(2004)]{str04} Strickland, D.~K., Heckman,
%   T.~M., Colbert, E.~J.~M., Hoopes, C.~G., \& Weaver, K.~A.\ 2004,
%   \apjs, 151, 193
% \bibitem[Taylor et al.(2010)]{tay10} Taylor, E.~N., Franx, M.,
%   Glazebrook, K., Brinchmann, J., van der Wel, A., \& van Dokkum,
%   P.~G.\ 2010, \apj, 720, 723
\bibitem[Thompson et al.(2005)]{tho05} Thompson, T.~A., Quataert, E.,
  \& Murray, N.\ 2005, \apj, 630, 167
\bibitem[Tremonti et al.(2007)]{tre07} Tremonti, C.~A., Moustakas, J.,
  \& Diamond-Stanic, A.~M.\ 2007, \apjl, 663, L77
% %\bibitem[Tremonti et al.(2012)]{tre12} Tremonti, C.~A., et al.\ 2012,
% %  in prep
\bibitem[Trujillo et al.(2007)]{tru07} Trujillo, I., Conselice, C.~J.,
  Bundy, K., et al.\ 2007, \mnras, 382, 109
\bibitem[van Dokkum et al.(2008)]{van08} van Dokkum, P.~G., et
  al.\ 2008, \apjl, 677, L5
% \bibitem[van Dokkum et al.(2010)]{van10} van Dokkum, P.~G., et
%   al.\ 2010, \apj, 709, 1018
% \bibitem[Vogt et al.(1994)]{vog94} Vogt, S.~S., et al.\ 1994,
%   \procspie, 2198, 362
\bibitem[Veilleux et al.(2006)]{vei06} Veilleux, S., Kim, D.-C., Peng,
  C.~Y., et al.\ 2006, \apj, 643, 707
\bibitem[Weiner et al.(2009)]{wei09} Weiner, B.~J., et al.\ 2009,
  \apj, 692, 187
\bibitem[Werner et al.(2004)]{wer04} Werner, M.~W., Roellig, T.~L.,
  Low, F.~J., et al.\ 2004, \apjs, 154, 1
\bibitem[Wright et al.(2010)]{wri10} Wright, E.~L., Eisenhardt,
  P.~R.~M., Mainzer, A.~K., et al.\ 2010, \aj, 140, 1868
% \bibitem[Wuyts et al.(2010)]{wuy10} Wuyts, S., Cox, T.~J., Hayward,
%   C.~C., Franx, M., Hernquist, L., Hopkins, P.~F., Jonsson, P., \& van
%   Dokkum, P.~G.\ 2010, \apj, 722, 1666
% \bibitem[Wuyts et al.(2011)]{wuy11} Wuyts, S., et al.\ 2011,
%   arXiv:1107.0317
\bibitem[Wuyts et al.(2011)]{wuy11} Wuyts, S., F{\"o}rster Schreiber,
  N.~M., van der Wel, A., et al.\ 2011, \apj, 742, 96
% \bibitem[Zhang \& Thompson(2010)]{zha10} Zhang, D., \& Thompson,
%   T.~A.\ 2010, arXiv:1005.4691
% \bibitem[Zirm et al.(2007)]{zir07} Zirm, A.~W., et al.\ 2007, \apj,
%   656, 66
\end{thebibliography}

\clearpage

%\clearpage

% \clearpage
% 
% \begin{figure*}[!t]
% \begin{center}
% \includegraphics[angle=0,scale=.60]{j0905_hires_mg.ps}
% \caption{Keck/HIRES spectra.}
% \label{fig:hires}
% \end{center}
% \end{figure*}

%\clearpage

\end{document}